### 2011-06-02 docs/pubs/EIM-2011/intro.tex

In this paper we describe extensions to the Ecological Metadata
Language (EML) \cite{} and suppporting tools for enabling improved
discovery and integration of ecological data sets. Our work is based
on the Extensible Observations Ontology (OBOE) \cite{}, which
represents a generic observational model implemented in OWL-DL for
describing domain-specific observation and measurement types. Our
approach adds additional metadata in the form of semantic annotations
that link attributes within datasets to OBOE terms for describing the
implicit observation and measurement types found within data
sets. Semantic annotations are executable in the sense that they can
be used to convert a data set into a collection of observation and
measurement instances, providing a more uniform representation for
expressing queries and performing integration.  To support the
creation of annotations, we have extended the Morpho metadata editor
with a high-level user interface as well as the Metacat data catalog
for storing and querying annotations through a new Semantic Mediation
API. This API can also be used to perform basic data-level integration
tasks using our prior work on the EML Data Manager Library \cite{}.

>>>>

In this paper we describe extensions to the Ecological Metadata
Language (EML) \cite{Fegraus07} and suppporting tools for enabling
improved discovery and integration of ecological data sets. Our work
is based on the Extensible Observations Ontology (OBOE)
\cite{bowers08,madin07:_ontol_for_descr_and_synth}, which represents a
generic observational model implemented in OWL-DL for describing
domain-specific observation and measurement types. Our approach adds
additional metadata in the form of semantic annotations that link
attributes within datasets to OBOE terms for describing the implicit
observation and measurement types found within data sets. Semantic
annotations are executable in the sense that they can be used to
convert a data set into a collection of observation and measurement
instances, providing a more uniform representation for expressing
queries and performing integration.  To support the creation of
annotations, we have extended the Morpho metadata editor
\cite{metacat02:_manag_heter_ecolog_data_using_morph} with a
high-level user interface as well as the Metacat data catalog
\cite{berkley01:_metac} for storing and querying annotations through a
new Semantic Mediation API. This API can also be used to perform basic
data-level integration tasks using our prior work on the EML Data
Manager Library \cite{leinfelder10:_metad_driven_approac_to_loadin}.
